Quantum H theorem, variational entropy, and relaxation processes

### 2021-03-10 tex/classical.tex
Removed comments (pdf remains the same)

%\section{Classical}
%\subsection{Boltzmann transport equation}
%\subsection{H theorem and the Maxwell-Boltzmann distribution}
%Todo esto de arriba en la forma tradicional y muy resumido. Solo remarcando lo fundamental.
%\subsection{Out of equilibrium, non-homogeneous distributions}.
%Entrar a la generalizacion, ahi si de manera muy detallada y con amplitud discutida
%fisicamente, del tratamiento con las celdas. Es decir, la demostracion del teorema H
%clasico pero con el enfoque de funciones de distribucion no homogeneas y por tanto
%con el tratamiento de celdas.

%----------------------------------

@@

We now use the variational method to find the extremal of $\mathcal{H}'$,
consistent with the restrictions
(\ref{eq:cellrestrictions}) and (\ref{eq:micro}), together with the corresponding Lagrange
multipliers $\{\alpha_M\}$ and $\{\beta_M\}$. This yields:
%
\begin{eqnarray}\label{eq:deltaHpdeltafpj}
    \frac{\delta \mathcal{H}'}{\delta f'_J(\vec{v}')} & = & \sum_{M=1}^{K}\int
      \frac{\delta}{\delta f'_J(\vec{v}')}\left[
        f'_M(\vec{v})\ln f'_M(\vec{v})
        \right]
       \mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int
       	\frac{\delta f'_M(\vec{v})}{\delta f'_J(\vec{v}')}
      \mathrm{d}^3v\nonumber\\
    & & -\sum_{M=1}^{K}\beta_M\int\epsilon(\vec{v})
    	\frac{\delta f'_M(\vec{v})}{\delta f'_J(\vec{v}')}
      \mathrm{d}^3v \nonumber\\
    & = & \ln f'_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
\end{eqnarray}
%
Solving the last line for $f'_J(\vec{v}')$ renders
%
\begin{comment}
\begin{eqnarray}
    \ln f'_J(\vec{v}') & = & \alpha_J+\beta_J \epsilon(\vec{v}')
       -1\quad\Rightarrow\quad f_J(\vec{v}')\\
    & = & e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1}\nonumber\\
    & = & Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{eq:relacion1},
\end{eqnarray}
\end{comment}
\begin{equation}\label{eq:relacion1}
	f'_J(\vec{v}') = C\exp\left({\alpha_J+\beta_J \epsilon(\vec{v}') }\right),
\end{equation}
%
where $C$ is a constant. We notice that applying the variational procedure on
$\mathcal{H}'$ predicts that when the equilibrium is reached, the
distribution function of each cell has the form of the Maxwell-Boltzmann
distribution function, which is consistent with the local equilibrium assumption.

@@

%---------------------------------------
%The text below is not needed, also it may be wrong. $g(t_0)>0$ indicates $f(t_0)>\bar f$, but this does
%not necessarily imply excess of particles and/or energy. E.g., a particular combination of a
%few missing particles and a considerable energy excess might render the corresponding
% $g(t_0)$ to be greater than zero.
\begin{comment}
Cells with ${{\dot{g}}'}_I\phantom{.}\!^{+}(\vec{v},t)$, $I=1,\,\dots\,,L$, 
have an excess of particles and/or energy (relative to the mean value in equilibrium).
In contrast, cells with ${{\dot{g}}'}_J\phantom{.}\!^{-}(\vec{v},t)$, $J=1,\,\dots\,,P$,
have a scarcity of particles and/or energy
(relative to the mean value in equilibrium).
In addition, $\dot{g}^{+}_{J}$ represents the
change on the deviation on cells that have an excess of particles or energy
while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that
have missing particles or energy.
On the other hand, $g^{+}_{J}$  represents the value of the deviation on cells
that have an excess of particles or energy. In contrast, $g^{-}_{J}$ represents
the value of the deviation on cells that have missing particles or energy.
Also, on the one hand, $\dot{g}^{+}_{J}<0$ describes the loss of particles
and/or energy and so, $g^{+}_{J}>0$. On the other hand, $\dot{g}^{-}_{J}>0$
describes the gain of particles and/or energy and therefore $g^{-}_{J}<0$.

We sort the previous ideas in the following form
%
\begin{equation}\label{separacionclassical}
\begin{array}{rl}
  g^{+}_{J}=+|g^{+}_{J}|; & \dot{g}^{+}_{J}=-|\dot{g}^{+}_{J}|,\\
  g^{-}_{J}=-|g^{-}_{J}|; & \dot{g}^{-}_{J}=+|\dot{g}^{-}_{J}|,
 \end{array}
\end{equation}
%
and consequently, (\ref{classicalH3}) obtains the following form

We can observe that $\bar{f}(\vec{v})$ in (\ref{eq:classicalH4}) is always
positive. Also the deviation and its derivative are positive, then all the
expression is positive. However, the global sign makes the derivative of the
$\mathcal{H}'$ with respect to time is always less to zero. With this we proved that
$\frac{d\mathcal{H}'}{dt}<0$. If the system is in equilibrium,
$g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{d\mathcal{H}'}{dt}=0$.
\end{comment}
%-----------------------------
%
If the system's initial state is sufficiently close to equilibrium, it is
physically suitable
to assume that $\left|g'_M(\vec{v},t_0)\right|\to 0$ as $t\to\infty$ in a monotonous manner,
thus ${{\dot{g}}'}_I\phantom{.}\!^{+}(\vec{v},t)\leq0$ and
${{\dot{g}}'}_J\phantom{.}\!^{-}(\vec{v},t)>0$, for $t\geq t_0$. Consequently,
Eq.~(\ref{eq:classicalH3}) can be re-written as
%
\begin{equation}\label{eq:classicalH4}
    \frac{d\mathcal{H}'}{dt}=-\left[
      \sum_J^{L}\int\bar f(\vec{v})|g_J^{+}(\vec{v},t)|
        |\dot g_J^{+}(\vec{v},t)|\mathrm{d}^3v
      +\sum_J^{P}\int\bar f(\vec{v})|g_J^{-}(\vec{v},t)|
      |\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v 
    \right].
\end{equation}
%
Since every integrand in Eq.~(\ref{eq:classicalH4}) is positive,
for all $t$ and $\vec v$, and
$\left|g'_M(\vec{v},t_0)\right|\to 0$ as $t\to\infty$, it follows that
%
\begin{equation}\label{eq:dHpdtleq0}
 \frac{d\mathcal{H}'}{dt}\leq0.\qquad\qquad\textrm{QED.}
\end{equation}
%

@@

\begin{comment}
Joining both results, we can say the following statement:
Consider a classical gas in a total volume $V$ (divided in $K$ cells of equal
volume elements), total energy $E$ and total number of free particles $N$.
Consider also the system has inhomogeneities and suffers a relaxation process.
If we define the following functional
%
\begin{equation}\label{CH3}
   \mathcal{H}'(t)=\sum_{M=1}^{K}\int f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v,
\end{equation}
%
where $f_M(\vec{v},t)$ is the local distribution function of each cell in the
system, then in the first-order approximation
%
\begin{equation}
    \frac{d\mathcal{H}'}{dt} \leq 0.
\end{equation}
%
This statement will be the classical $H$-theorem with inhomogeneities.

In the next section, we present a proposal quantum version of the
$H$-theorem defined by Tolman and our proposal quantum version of the $H$
theorem applying the method of the volume divided into cells.
\end{comment}

>>>>

@@

We now use the variational method to find the extremal of $\mathcal{H}'$,
consistent with the restrictions
(\ref{eq:cellrestrictions}) and (\ref{eq:micro}), together with the corresponding Lagrange
multipliers $\{\alpha_M\}$ and $\{\beta_M\}$. This yields:
%
\begin{eqnarray}\label{eq:deltaHpdeltafpj}
    \frac{\delta \mathcal{H}'}{\delta f'_J(\vec{v}')} & = & \sum_{M=1}^{K}\int
      \frac{\delta}{\delta f'_J(\vec{v}')}\left[
        f'_M(\vec{v})\ln f'_M(\vec{v})
        \right]
       \mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int
       	\frac{\delta f'_M(\vec{v})}{\delta f'_J(\vec{v}')}
      \mathrm{d}^3v\nonumber\\
    & & -\sum_{M=1}^{K}\beta_M\int\epsilon(\vec{v})
    	\frac{\delta f'_M(\vec{v})}{\delta f'_J(\vec{v}')}
      \mathrm{d}^3v \nonumber\\
    & = & \ln f'_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.
\end{eqnarray}
%
Solving the last line for $f'_J(\vec{v}')$ renders
%
\begin{equation}\label{eq:relacion1}
	f'_J(\vec{v}') = C\exp\left({\alpha_J+\beta_J \epsilon(\vec{v}') }\right),
\end{equation}
%
where $C$ is a constant. We notice that applying the variational procedure on
$\mathcal{H}'$ predicts that when the equilibrium is reached, the
distribution function of each cell has the form of the Maxwell-Boltzmann
distribution function, which is consistent with the local equilibrium assumption.

@@

If the system's initial state is sufficiently close to equilibrium, it is
physically suitable
to assume that $\left|g'_M(\vec{v},t_0)\right|\to 0$ as $t\to\infty$ in a monotonous manner,
thus ${{\dot{g}}'}_I\phantom{.}\!^{+}(\vec{v},t)\leq0$ and
${{\dot{g}}'}_J\phantom{.}\!^{-}(\vec{v},t)>0$, for $t\geq t_0$. Consequently,
Eq.~(\ref{eq:classicalH3}) can be re-written as
\begin{equation}\label{eq:classicalH4}
    \frac{d\mathcal{H}'}{dt}=-\left[
      \sum_J^{L}\int\bar f(\vec{v})|g_J^{+}(\vec{v},t)|
        |\dot g_J^{+}(\vec{v},t)|\mathrm{d}^3v
      +\sum_J^{P}\int\bar f(\vec{v})|g_J^{-}(\vec{v},t)|
      |\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v 
    \right].
Since every integrand in Eq.~(\ref{eq:classicalH4}) is positive,
for all $t$ and $\vec v$, and
$\left|g'_M(\vec{v},t_0)\right|\to 0$ as $t\to\infty$, it follows that
\begin{equation}\label{eq:dHpdtleq0}
 \frac{d\mathcal{H}'}{dt}\leq0.\qquad\qquad\textrm{QED.}

@@
